---
title: "Getting Started with baseballr"
description: "Getting started with using <code>baseballr</code> and men's baseball data."
author: "Saiem Gilani <br> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>"
date: "`r Sys.Date()`"
opengraph:
  image: 
    src: "https://raw.githubusercontent.com/billpetti/baseballr/master/logo.png"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with baseballr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# You can install using the pacman package using the following code:
# if (!requireNamespace('pacman', quietly = TRUE)){
#   install.packages('pacman')
# }
# pacman::p_load_current_gh("r-lib/pkgapi")
# library(pkgapi)
# pkg <- pkgapi::map_package(path = "../")
# library(dplyr)
# function_calls <- pkg$calls
# exported <- pkg$defs %>% 
#   dplyr::filter(exported == TRUE) %>% 
#   dplyr::select(all_of(c("name", "file"))) 
# 
# mlb_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^mlb_*"))
# 
# bref_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^bref_*"))
# fg_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^fg_*"))
# sc_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^statcast_*"))
# ncaa_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^ncaa_*"))
# chadwick_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^chadwick_*|^playerid_*|^playername_*"))
# retrosheet_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^retrosheet_*"))
old <- options(rmarkdown.html_vignette.check_title = FALSE)
pkg_name <- "billpetti/baseballr"
url <- paste0("https://raw.githubusercontent.com/", pkg_name, "/master/DESCRIPTION")
x <- readLines(url)
remote_version <- gsub("Version:\\s*", "", x[grep('Version:', x)])
```

Welcome folks,

I'm Saiem Gilani, one of the [authors](https://billpetti.github.io/baseballr/authors.html "Authors and contributors to baseballr") of [`baseballr`](https://billpetti.github.io/baseballr/), and I hope to give the community a high-quality resource for accessing men's baseball data for statistical analysis, baseball research, and more. I am excited to show you some of what you can do with this edition of the package.

### **Installing R and RStudio**

1.  Head to <https://cran.r-project.org>
2.  Select the appropriate link for your operating system (Windows, Mac OS X, or Linux)

-   **Windows** - Select base and download the most recent version
-   **Mac OS X** - Select Latest Release, but check to make sure your OS is the correct version. Look through Binaries for Legacy OS X Systems if you are on an older release
-   **Linux** - Select the appropriate distro and follow the installation instructions

3.  Head to [RStudio.com](https://posit.co/download/rstudio-desktop/#download "Download the appropriate version of RStudio (Free) for your operating system to use with R")
4.  Follow the associated download and installation instructions for RStudio.
5.  Start peering over the [RStudio IDE Cheatsheet](https://raw.githubusercontent.com/rstudio/cheatsheets/main/rstudio-ide.pdf). *An IDE is an integrated development environment.*
6.  For **Windows** users: I recommend you install [Rtools](https://cran.r-project.org/bin/windows/Rtools/). This is not an R package! It is “a collection of resources for building packages for R under Microsoft Windows, or for building R itself”. Go to [https://cran.r-project.org/bin/windows/Rtools/](https://cran.r-project.org/bin/windows/Rtools/) and follow the directions for installation.




## **Install** [**`baseballr`**](https://billpetti.github.io/baseballr/)
```{r install_baseballr_gs, message=FALSE,eval=FALSE}
# You can install using the pacman package using the following code:
if (!requireNamespace('pacman', quietly = TRUE)){
  install.packages('pacman')
}
pacman::p_load_current_gh("billpetti/baseballr")
```

### **The Data**

There are generally speaking **eight** men's baseball data sources accessible from this package:

-   [`baseballr-data` repo](https://github.com/sportsdataverse/baseballr-data) (@SportsDataverse)
-   [MLB Stats API](https://www.mlb.com/)
-   [Baseball Savant's Statcast](https://baseballsavant.mlb.com/)
-   [Chadwick Bureau's Public Register of Baseball Players](https://github.com/chadwickbureau/register/)
-   [Baseball Reference](https://www.baseball-reference.com/)
-   [FanGraphs](https://www.fangraphs.com/)
-   [Retrosheet](https://www.retrosheet.org/)
-   [NCAA](https://stats.ncaa.org/)

#### **Function names indicate the data source**

As of [baseballr v1.0.0](https://billpetti.github.io/baseballr/news/index.html#baseballr-1.0.0), a function naming convention was implemented to have the data source indicator appear at the start of the function name:

-   Functions that use the [`baseballr-data` repository](https://github.com/sportsdataverse/baseballr-data) will contain `load_` or `update_` in the function name and would be considered loading functions for the play-by-play data, team box scores, and player box scores.

-   Functions that use the MLB Stats API start with `mlb_` by convention and should be assumed as `get` functions. As of `baseballr` version `r remote_version`, the package exports ~88 functions covering the MLB Stats API.

-   Functions that use one of [Baseball Savant's Statcast](https://baseballsavant.mlb.com/) APIs start with `statcast_` by convention and should be assumed as `get` functions. These functions allow for live access to Statcast data for the MLB games in-progress. As of `baseballr` version `r remote_version`, the package exports ~5 Statcast-related functions.

-   Functions that use [Chadwick Bureau's Public Register of Baseball Players](https://github.com/chadwickbureau/register/) start with `chadwick_`, `playerid_`, or `playername_` by convention and should be assumed as `get` functions. These functions allow for access to the Bureau's public register of baseball players. As of `baseballr` version `r remote_version`, the package exports 3 functions sourced using the Chadwick Bureau's public register of baseball players.

-   Functions that use Baseball Reference's website start with `bref_` by convention and should be assumed as `get` functions. As of `baseballr` version `r remote_version`, the package exports ~4 functions covering [Baseball Reference](https://www.baseball-reference.com/).

-   Functions that use FanGraphs's baseball website start with `fg_` by convention and should be assumed as `get` functions. As of `baseballr` version `r remote_version`, the package exports ~11 functions covering [FanGraphs.com](https://www.fangraphs.com/).

-   Functions that use Retrosheet's baseball data start with `retrosheet_` by convention and should be assumed as `get` functions. As of `baseballr` version `r remote_version`, the package exports 1 function for [Retrosheet Data](https://www.retrosheet.org/).

-   Functions that use the NCAA website start with `ncaa_` by convention and should be assumed as `get` functions. As of `baseballr` version `r remote_version`, the package exports ~8 function covering the [NCAA Stats portal](https://stats.ncaa.org/).


## Follow the SportsDataverse (@SportsDataverse) on Twitter and star this repo

<!-- [![Twitter Follow](https://img.shields.io/twitter/follow/SportsDataverse?color=blue&label=%40SportsDataverse&logo=twitter&style=for-the-badge)](https://twitter.com/SportsDataverse)  -->

[![GitHub stars](https://img.shields.io/github/stars/billpetti/baseballr.svg?color=eee&logo=github&style=for-the-badge&label=Star%20baseballr&maxAge=2592000)](https://github.com/billpetti/baseballr/stargazers/)

## **Our Authors**

-   Bill Petti (@BillPetti)       
<!-- <a href="https://twitter.com/BillPetti" target="blank"><img src="https://img.shields.io/twitter/follow/BillPetti?color=blue&label=%40BillPetti&logo=twitter&style=for-the-badge" alt="@BillPetti" /></a> -->
<a href="https://github.com/BillPetti" target="blank"><img src="https://img.shields.io/github/followers/BillPetti?color=eee&logo=Github&style=for-the-badge" alt="@BillPetti" /></a>

-   Saiem Gilani (@saiemgilani)       
<!-- <a href="https://twitter.com/saiemgilani" target="blank"><img src="https://img.shields.io/twitter/follow/saiemgilani?color=blue&label=%40saiemgilani&logo=twitter&style=for-the-badge" alt="@saiemgilani" /></a> -->
<a href="https://github.com/saiemgilani" target="blank"><img src="https://img.shields.io/github/followers/saiemgilani?color=eee&logo=Github&style=for-the-badge" alt="@saiemgilani" /></a>