---
title: "Getting Started with baseballr"
description: "Getting started with using <code>baseballr</code> and men's baseball data."
author: "Saiem Gilani <br><a href='https://twitter.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=twitter&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>"
date: "`r Sys.Date()`"
opengraph:
  image: 
    src: "https://github.com/saiemgilani/hoopR-data/blob/master/themes/hoopR_gh.png?raw=true"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with baseballr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# You can install using the pacman package using the following code:
# if (!requireNamespace('pacman', quietly = TRUE)){
#   install.packages('pacman')
# }
# pacman::p_load_current_gh("r-lib/pkgapi")
# library(pkgapi)
# pkg <- pkgapi::map_package(path = "../")
# library(dplyr)
# function_calls <- pkg$calls
# exported <- pkg$defs %>% 
#   dplyr::filter(exported == TRUE) %>% 
#   dplyr::select(all_of(c("name", "file"))) 
# 
# mlb_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^mlb_*"))
# 
# bref_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^bref_*"))
# fg_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^fg_*"))
# sc_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^statcast_*"))
# ncaa_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^ncaa_*"))
# chadwick_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^chadwick_*|^playerid_*|^playername_*"))
# retrosheet_functions <- exported %>% 
#   dplyr::filter(stringr::str_detect(.data$name,pattern = "^retrosheet_*"))
old <- options(rmarkdown.html_vignette.check_title = FALSE)
pkg_name <- "billpetti/baseballr"
url <- paste0("https://raw.githubusercontent.com/", pkg_name, "/master/DESCRIPTION")
x <- readLines(url)
remote_version <- gsub("Version:\\s*", "", x[grep('Version:', x)])
```

Welcome folks,

I'm Saiem Gilani, one of the [authors](https://billpetti.github.io/baseballr/authors.html "Authors and contributors to baseballr") of [`baseballr`](https://billpetti.github.io/baseballr/), and I hope to give the community a high-quality resource for accessing men's baseball data for statistical analysis, baseball research, and more. I am excited to show you some of what you can do with this edition of the package.

## **Install** [**`baseballr`**](https://billpetti.github.io/baseballr/)
```{r install_baseballr_gs, message=FALSE,eval=FALSE}
# You can install using the pacman package using the following code:
if (!requireNamespace('pacman', quietly = TRUE)){
  install.packages('pacman')
}
pacman::p_load_current_gh("billpetti/baseballr")
```

### **The Data**

There are generally speaking **eight** men's baseball data sources accessed from this package:

-   [`baseballr-data` repo](https://github.com/saiemgilani/baseballr-data) [![Twitter Follow](https://img.shields.io/twitter/follow/sportsdataverse?color=blue&label=%40sportsdataverse&logo=twitter&style=for-the-badge)](https://twitter.com/sportsdataverse)
-   [MLB Stats API](https://www.mlb.com/)
-   [Baseball Savant's Statcast](https://baseballsavant.mlb.com/)
-   [Chadwick Bureau's Public Register of Baseball Players](https://github.com/chadwickbureau/register/)
-   [Baseball Reference](https://www.baseball-reference.com/)
-   [FanGraphs](https://www.fangraphs.com/)
-   [Retrosheet](https://www.retrosheet.org/)
-   [NCAA](https://stats.ncaa.org/)

#### **Function names indicate the data source**

-   Functions that use the [`baseballr-data` repository](https://github.com/saiemgilani/baseballr-data) will contain `load_` or `update_` in the function name and would be considered loading functions for the play-by-play data, team box scores, and player box scores.

-   Functions that use the MLB Stats API start with `mlb_` by convention and should be assumed as `get` functions. As of `baseballr` version `r remote_version`, the package exports ~88 functions covering the MLB Stats API

-   Functions that use one of [Baseball Savant's Statcast](https://baseballsavant.mlb.com/) APIs start with `statcast_` by convention and should be assumed as `get` functions. These functions allow for live access to Statcast data for the MLB games in-progress. As of `baseballr` version `r remote_version`, the package exports ~5 Statcast-related functions.

-   Functions that use [Chadwick Bureau's Public Register of Baseball Players](https://github.com/chadwickbureau/register/) start with `chadwick_`, `playerid_`, or `playername_` by convention and should be assumed as `get` functions. These functions allow for access to the Bureau's public register of baseball players. As of `baseballr` version `r remote_version`, the package exports 3 functions sourced using the Chadwick Bureau's public register of baseball players.

-   Functions that use Baseball Reference's website start with `bref_` by convention and should be assumed as `get` functions. As of `baseballr` version `r remote_version`, the package exports ~4 functions covering [Baseball Reference](https://www.baseball-reference.com/).

-   Functions that use FanGraphs's baseball website start with `fg_` by convention and should be assumed as `get` functions. As of `baseballr` version `r remote_version`, the package exports ~11 functions covering [FanGraphs.com](https://www.fangraphs.com/).

-   Functions that use Retrosheet's baseball data start with `retrosheet_` by convention and should be assumed as `get` functions. As of `baseballr` version `r remote_version`, the package exports 1 function for [Retrosheet Data](https://www.retrosheet.org/).

-   Functions that use the NCAA website start with `ncaa_` by convention and should be assumed as `get` functions. As of `baseballr` version `r remote_version`, the package exports ~8 function covering the [NCAA Stats portal](https://stats.ncaa.org/).

## Follow the [SportsDataverse](https://twitter.com/SportsDataverse) on Twitter and star this repo

[![Twitter Follow](https://img.shields.io/twitter/follow/SportsDataverse?color=blue&label=%40SportsDataverse&logo=twitter&style=for-the-badge)](https://twitter.com/SportsDataverse) 

[![GitHub stars](https://img.shields.io/github/stars/billpetti/baseballr.svg?color=eee&logo=github&style=for-the-badge&label=Star%20baseballr&maxAge=2592000)](https://github.com/billpetti/baseballr/stargazers/)

## **Our Authors**

-   [Bill Petti](https://twitter.com/BillPetti)       
<a href="https://twitter.com/BillPetti" target="blank"><img src="https://img.shields.io/twitter/follow/BillPetti?color=blue&label=%40BillPetti&logo=twitter&style=for-the-badge" alt="@BillPetti" /></a>
<a href="https://github.com/BillPetti" target="blank"><img src="https://img.shields.io/github/followers/BillPetti?color=eee&logo=Github&style=for-the-badge" alt="@BillPetti" /></a>

-   [Saiem Gilani](https://twitter.com/saiemgilani)       
<a href="https://twitter.com/saiemgilani" target="blank"><img src="https://img.shields.io/twitter/follow/saiemgilani?color=blue&label=%40saiemgilani&logo=twitter&style=for-the-badge" alt="@saiemgilani" /></a>
<a href="https://github.com/saiemgilani" target="blank"><img src="https://img.shields.io/github/followers/saiemgilani?color=eee&logo=Github&style=for-the-badge" alt="@saiemgilani" /></a>
