---
title: "NCAA Scraping"
author: "Bill Petti, Saiem Gilani <br><a href='https://x.com/BillPetti' target='blank'><img src='https://img.shields.io/twitter/follow/BillPetti?color=blue&amp;label=%40BillPetti&amp;logo=x&amp;style=for-the-badge' alt='@BillPetti'/></a> <a href='https://github.com/BillPetti' target='blank'><img src='https://img.shields.io/github/followers/BillPetti?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@BillPetti'/></a> <a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>"
date: "2016-11-22"
output: html_document
description: "Pulling college baseball player and team statistics from the NCAA's website with <code>ncaa_scrape()</code>, including batting, pitching and fielding splits by season."
opengraph:
  image:
    src: "https://raw.githubusercontent.com/sportsdataverse/baseballr-data/main/themes/baseballr_gh.png"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The latest release of the [`baseballr`](https://billpetti.github.io/baseballr/) includes a function for acquiring player statistics from the [NCAA's website](http://stats.ncaa.org) for baseball teams across the three major divisions (I, II, III).

In order to look up teams, you can either load the teams for all divisions from the `baseballr-data` repository or access them directly from the NCAA website for a given year and division.

Loading from the baseballr-data repository:
```{r ncaa_teams}
library(baseballr)
library(dplyr)
ncaa_teams_df <- load_ncaa_baseball_teams()
```
From the NCAA website:
```{r ncaa_teams_website}
try(ncaa_teams(year = most_recent_ncaa_baseball_season(), division = "1"))
```

The function, ```ncaa_team_player_stats()```, requires the user to pass values for three parameters for the function to work:

`team_id`: numerical code used by the NCAA for each school
`year`: a four-digit year
`type`: whether to pull data for batters or pitchers

If you want to pull batting statistics for Florida State for the `r baseballr::most_recent_ncaa_baseball_season()` season, you would use the following:

```{r ncaa_scrape_batting}

team_id <- ncaa_teams_df %>% 
  dplyr::filter(.data$team_name == "Florida St.") %>% 
  dplyr::select("team_id") %>% 
  dplyr::distinct() %>% 
  dplyr::pull("team_id")

year <- most_recent_ncaa_baseball_season()

ncaa_team_player_stats(team_id = team_id, year = year, "batting")

```


The same can be done for pitching, just by changing the `type` parameter:

```{r ncaa_scrape_pitching}
ncaa_team_player_stats(team_id = team_id, year = year,  "pitching")
```


Now, the function is dependent on the user knowing the `team_id` used by the NCAA website. Given that, I've included a `ncaa_school_id_lu` function so that users can find the `team_id` they need.

Just pass a string to the function and it will return possible matches based on the school's name:

```{r ncaa_sch_id_lu}
ncaa_school_id_lu("Vand")
```

# **Our Authors**
-   [Bill Petti](https://x.com/BillPetti)
<a href='https://x.com/BillPetti' target='blank'><img src='https://img.shields.io/twitter/follow/BillPetti?color=blue&amp;label=%40BillPetti&amp;logo=x&amp;style=for-the-badge' alt='@BillPetti'/></a>
<a href='https://github.com/BillPetti' target='blank'><img src='https://img.shields.io/github/followers/BillPetti?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@BillPetti'/></a>
-   [Saiem Gilani](https://x.com/saiemgilani)
<a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a>
<a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>

## **Our Contributors**
-   [Ben Baumer](https://x.com/BaumerBen)
<a href='https://x.com/BaumerBen' target='blank'><img src='https://img.shields.io/twitter/follow/BaumerBen?color=blue&amp;label=%40BaumerBen&amp;logo=x&amp;style=for-the-badge' alt='@BaumerBen'/></a>
-   [Ben Dilday](https://x.com/BenDilday)
<a href='https://x.com/BenDilday' target='blank'><img src='https://img.shields.io/twitter/follow/BenDilday?color=blue&amp;label=%40BenDilday&amp;logo=x&amp;style=for-the-badge' alt='@BenDilday'/></a>
-   [Robert Frey](https://x.com/RobertFrey40)
<a href='https://x.com/RobertFrey40' target='blank'><img src='https://img.shields.io/twitter/follow/RobertFrey40?color=blue&amp;label=%40RobertFrey40&amp;logo=x&amp;style=for-the-badge' alt='@RobertFrey40'/></a>
-   [Camden Kay](https://x.com/k_camden)
<a href='https://x.com/k_camden' target='blank'><img src='https://img.shields.io/twitter/follow/k_camden?color=blue&amp;label=%40k_camden&amp;logo=x&amp;style=for-the-badge' alt='@k_camden'/></a>

## **Citation**

To cite the [**`baseballr`**](https://baseballr.sportsdataverse.org/) R package in publications, use:

BibTex Citation

```bibtex
@misc{baseballr,
  author = {Bill Petti, Saiem Gilani},
  title = {baseballr: An R Package for Baseball Data Acquisition and Analysis},
  url = {https://baseballr.sportsdataverse.org/},
  year = {2026}
}
```

## **Related SportsDataverse packages**

-   [**cfbfastR**](https://cfbfastR.sportsdataverse.org/) - college football
-   [**hoopR**](https://hoopR.sportsdataverse.org/) - men's basketball
-   [**wehoop**](https://wehoop.sportsdataverse.org/) - women's basketball
-   [**baseballr**](https://baseballr.sportsdataverse.org/) - baseball
-   [**fastRhockey**](https://fastRhockey.sportsdataverse.org/) - hockey
-   [**oddsapiR**](https://oddsapiR.sportsdataverse.org/) - betting odds
-   [**sportyR**](https://sportyR.sportsdataverse.org/) - playing surfaces
-   [**sportsdataverse-py**](https://py.sportsdataverse.org/) - the Python package
-   [**sportsdataverse-R**](https://r.sportsdataverse.org/) - the R meta-package
