---
title: "NCAA Scraping"
author: "Bill Petti"
date: "2016-11-22"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NCAA Scraping}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The latest release of the [`baseballr`](https://billpetti.github.io/baseballr/) includes a function for acquiring player statistics from the [NCAA's website](http://stats.ncaa.org) for baseball teams across the three major divisions (I, II, III).

In order to look up teams, you can either load the teams for all divisions from the `baseballr-data` repository or access them directly from the NCAA website for a given year and division.

Loading from the baseballr-data repository:
```{r ncaa_teams}
ncaa_teams_df <- load_ncaa_baseball_teams()
```
From the NCAA website:
```{r ncaa_teams_website}
ncaa_teams(year = most_recent_ncaa_baseball_season(), division = "1")
```

The function, `ncaa_scrape`, requires the user to pass values for three parameters for the function to work:

`team_id`: numerical code used by the NCAA for each school
`year`: a four-digit year
`type`: whether to pull data for batters or pitchers

If you want to pull batting statistics for Florida State for the `r baseballr:::most_recent_ncaa_baseball_season()` season, you would use the following:

```{r ncaa_scrape_batting}
library(baseballr)
library(dplyr)
teamid <- ncaa_teams_df %>% 
  dplyr::filter(.data$team_name == "Florida St.") %>% 
  dplyr::select("team_id") %>% 
  dplyr::distinct() %>% 
  dplyr::pull("team_id")

year <- most_recent_ncaa_baseball_season()
ncaa_scrape(teamid = teamid, year = year, "batting")

```


The same can be done for pitching, just by changing the `type` parameter:

```{r ncaa_scrape_pitching}
ncaa_scrape(teamid = teamid, year = year,  "pitching")
```


Now, the function is dependent on the user knowing the `school_id` used by the NCAA website. Given that, I've included a `ncaa_school_id_lu` function so that users can find the `school_id` they need.

Just pass a string to the function and it will return possible matches based on the school's name:

```{r ncaa_sch_id_lu}
ncaa_school_id_lu("Vand")
```

