---
title: "NCAA Scraping"
author: "Bill Petti"
date: "2016-11-22"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NCAA Scraping}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The latest release of the [`baseballr`](https://billpetti.github.io/baseballr/) includes a function for acquiring player statistics from the [NCAA's website](http://stats.ncaa.org) for baseball teams across the three major divisions (I, II, III).

The function, `ncaa_scrape`, requires the user to pass values for three parameters for the function to work:

`school_id`: numerical code used by the NCAA for each school
`year`: a four-digit year
`type`: whether to pull data for batters or pitchers

If you want to pull batting statistics for Vanderbilt for the 2013 season, you would use the following:

```{r}
library(baseballr)
library(dplyr)
ncaa_scrape(736, 2013, "batting") %>%
  select(year:OBPct)
```


The same can be done for pitching, just by changing the `type` parameter:

```{r}
ncaa_scrape(736, 2013, "pitching") %>%
  select(year:ERA)
```


Now, the function is dependent on the user knowing the `school_id` used by the NCAA website. Given that, I've included a `ncaa_school_id_lu` function so that users can find the `school_id` they need.

Just pass a string to the function and it will return possible matches based on the school's name:

```{r}
ncaa_school_id_lu("Vand")
```

