---
title: "Plotting Statcast data"
author: "Bill Petti"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting Statcast data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

In this example, the `baseballr` package is used to acquire Statcast data for Mookie Betts from 2015-2016.

The data is then processed and plotted to show how his launch angle and batted ball speed have changed from year to year

```{r}
library(baseballr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(zoo)
```

## find Mookie Betts' MLBAMID

```{r}
betts_id <- playerid_lookup("Betts") %>%
  filter(first_name == "Mookie") %>%
  select(mlbam_id, first_name, last_name)
```

## scrape Betts' Statcast data, by pitch removing those with a batted ball speed of 0

```{r}
betts_15 <- scrape_statcast_savant("2015-03-31", "2015-10-31", playerid = betts_id[1,1], player_type = 'batter')
betts_16 <- scrape_statcast_savant("2016-03-31", "2016-10-31", playerid = betts_id[1,1], player_type = 'batter')   
betts <- bind_rows(betts_15, betts_16) 3%>%
  mutate(Year = as.factor(substr(game_date,1,4))) %>%
  filter(type == "X") %>%
  filter(launch_speed != 0)
```

## calculate average launch angles and batted ball speeds by game

```{r}
betts_grpd <- betts %>%
  group_by(game_date) %>%
  summarise(`Average Launch Angle` = mean(launch_angle, na.rm = TRUE), 
            `Average Batted Ball Speed` = mean(launch_speed, na.rm = TRUE)) %>%
  ungroup() %>%
  melt(id=c("game_date")) %>%
  mutate(Year = as.factor(substr(game_date,1,4)))
```

## calculate Betts' average launch angle and batted ball speed by year

```{r}
betts_avg_speed_yr <- betts %>%
  group_by(Year) %>%
  summarise(speed = round(mean(launch_speed, na.rm = TRUE),1), angle = round(mean(launch_angle, na.rm = TRUE),1))
```

## plot the data

```{r}
betts_grpd %>%
  ggplot(aes(game_date, value)) +
  geom_point() +
  stat_smooth(aes(group = Year, color = Year)) +
  facet_wrap(~variable, scales = "free_y") + 
  ggtitle("\nMookie Betts: 2015 vs. 2016\n") + 
  labs(subtitle = paste0("Betts has lowered his launch angle in 2016, from ", betts_avg_speed_yr[1,3], " degrees in 2015 to ", betts_avg_speed_yr[2,3], " degrees this year.\n\n"), 
       caption = "@BillPetti\nData from baseballsavant.mlb.com\nData acquired with the baseballr package") +
  ylab("Angle = Degrees, Speed = MPH\n") +
  xlab("\nDate") +
  #theme_bp_grey() + 
  theme(legend.position = "bottom", strip.text.x = element_text(face = "bold", size = 14), plot.subtitle = element_text(hjust=-.12)) +
  scale_color_manual(values = c("#5F9ED1", "#FF800E"))
```

## export plot to your working directory

```{r, eval=FALSE}
ggsave("betts_angle_speed_year.png", scale = 1.2, width = 14, height = 8.5, units = "in")
```
