---
title: "Plotting Statcast data"
author: "Bill Petti, Saiem Gilani <br><a href='https://x.com/BillPetti' target='blank'><img src='https://img.shields.io/twitter/follow/BillPetti?color=blue&amp;label=%40BillPetti&amp;logo=x&amp;style=for-the-badge' alt='@BillPetti'/></a> <a href='https://github.com/BillPetti' target='blank'><img src='https://img.shields.io/github/followers/BillPetti?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@BillPetti'/></a> <a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>"
date: "`r Sys.Date()`"
output: html_document
description: "Drawing strike-zone plots and pitch locations from Statcast data with <code>ggplot2</code>, including how to place the zone correctly for left- and right-handed batters."
opengraph:
  image:
    src: "https://raw.githubusercontent.com/sportsdataverse/baseballr-data/main/themes/baseballr_gh.png"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

In this example, the `baseballr` package is used to acquire Statcast data for Mookie Betts from 2015-2016.

The data is then processed and plotted to show how his launch angle and batted ball speed have changed from year to year

```{r}
library(baseballr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(zoo)
```

## find Mookie Betts' MLBAMID

```{r}
betts_id <- playerid_lookup("Betts") %>%
  dplyr::filter(first_name == "Mookie") %>%
  dplyr::select(mlbam_id, first_name, last_name)
```

## scrape Betts' Statcast data, by pitch removing those with a batted ball speed of 0

```{r}
betts_15 <- statcast_search("2015-03-31", "2015-10-31", playerid = betts_id[1,1], player_type = 'batter')
betts_16 <- statcast_search("2016-03-31", "2016-10-31", playerid = betts_id[1,1], player_type = 'batter')   
betts <- dplyr::bind_rows(betts_15, betts_16) %>%
  dplyr::mutate(Year = as.factor(substr(game_date,1,4))) %>%
  dplyr::filter(type == "X") %>%
  dplyr::filter(launch_speed != 0)
```

## calculate average launch angles and batted ball speeds by game

```{r}
betts_grpd <- betts %>%
  dplyr::group_by(game_date) %>%
  dplyr::summarise(
    `Average Launch Angle` = mean(launch_angle, na.rm = TRUE), 
    `Average Batted Ball Speed` = mean(launch_speed, na.rm = TRUE)) %>%
  dplyr::ungroup() %>%
  melt(id=c("game_date")) %>%
  dplyr::mutate(Year = as.factor(substr(game_date,1,4)))
```

## calculate Betts' average launch angle and batted ball speed by year

```{r}
betts_avg_speed_yr <- betts %>%
  dplyr::group_by(Year) %>%
  dplyr::summarise(
    speed = round(mean(launch_speed, na.rm = TRUE),1), 
    angle = round(mean(launch_angle, na.rm = TRUE),1))
```

## plot the data

```{r}
betts_grpd %>%
  ggplot(aes(game_date, value)) +
  geom_point() +
  stat_smooth(aes(group = Year, color = Year)) +
  facet_wrap(~variable, scales = "free_y") + 
  ggtitle("\nMookie Betts: 2015 vs. 2016\n") + 
  labs(subtitle = paste0("Betts has lowered his launch angle in 2016, from ", betts_avg_speed_yr[1,3], " degrees in 2015 to ", betts_avg_speed_yr[2,3], " degrees this year.\n\n"), 
       caption = "@BillPetti\nData from baseballsavant.mlb.com\nData acquired with the baseballr package") +
  ylab("Angle = Degrees, Speed = MPH\n") +
  xlab("\nDate") +
  #theme_bp_grey() + 
  theme(legend.position = "bottom", strip.text.x = element_text(face = "bold", size = 14), plot.subtitle = element_text(hjust=-.12)) +
  scale_color_manual(values = c("#5F9ED1", "#FF800E"))
```

## Uncomment and run to export plot to your working directory

```{r, eval=FALSE}
# ggsave("betts_angle_speed_year.png", scale = 1.2, width = 14, height = 8.5, units = "in")
```

## Coloring by pitch type

`statcast_pitch_colors()` returns Baseball Savant's own pitch-type palette, so
plots colored by `pitch_type` match what readers see on Savant:

```{r, eval=FALSE}
pal <- statcast_pitch_colors()
pitch_pal <- setNames(pal$color, pal$pitch_type)

betts_16 |>
  dplyr::filter(!is.na(pitch_type)) |>
  ggplot(aes(x = plate_x, y = plate_z, color = pitch_type)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = pitch_pal) +
  coord_fixed() +
  theme_bw()
```

## Pitch locations with the strike zone

`ggpitchzone()` plots plate-crossing locations from the catcher's perspective
with the batter's strike zone overlaid, using the Savant palette by default:

```{r, eval=FALSE}
ggpitchzone(betts_16)
```

# **Our Authors**
-   [Bill Petti](https://x.com/BillPetti)
<a href='https://x.com/BillPetti' target='blank'><img src='https://img.shields.io/twitter/follow/BillPetti?color=blue&amp;label=%40BillPetti&amp;logo=x&amp;style=for-the-badge' alt='@BillPetti'/></a>
<a href='https://github.com/BillPetti' target='blank'><img src='https://img.shields.io/github/followers/BillPetti?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@BillPetti'/></a>
-   [Saiem Gilani](https://x.com/saiemgilani)
<a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a>
<a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>

## **Our Contributors**
-   [Ben Baumer](https://x.com/BaumerBen)
<a href='https://x.com/BaumerBen' target='blank'><img src='https://img.shields.io/twitter/follow/BaumerBen?color=blue&amp;label=%40BaumerBen&amp;logo=x&amp;style=for-the-badge' alt='@BaumerBen'/></a>
-   [Ben Dilday](https://x.com/BenDilday)
<a href='https://x.com/BenDilday' target='blank'><img src='https://img.shields.io/twitter/follow/BenDilday?color=blue&amp;label=%40BenDilday&amp;logo=x&amp;style=for-the-badge' alt='@BenDilday'/></a>
-   [Robert Frey](https://x.com/RobertFrey40)
<a href='https://x.com/RobertFrey40' target='blank'><img src='https://img.shields.io/twitter/follow/RobertFrey40?color=blue&amp;label=%40RobertFrey40&amp;logo=x&amp;style=for-the-badge' alt='@RobertFrey40'/></a>
-   [Camden Kay](https://x.com/k_camden)
<a href='https://x.com/k_camden' target='blank'><img src='https://img.shields.io/twitter/follow/k_camden?color=blue&amp;label=%40k_camden&amp;logo=x&amp;style=for-the-badge' alt='@k_camden'/></a>

## **Citation**

To cite the [**`baseballr`**](https://baseballr.sportsdataverse.org/) R package in publications, use:

BibTex Citation

```bibtex
@misc{baseballr,
  author = {Bill Petti, Saiem Gilani},
  title = {baseballr: An R Package for Baseball Data Acquisition and Analysis},
  url = {https://baseballr.sportsdataverse.org/},
  year = {2026}
}
```

## **Related SportsDataverse packages**

-   [**cfbfastR**](https://cfbfastR.sportsdataverse.org/) - college football
-   [**hoopR**](https://hoopR.sportsdataverse.org/) - men's basketball
-   [**wehoop**](https://wehoop.sportsdataverse.org/) - women's basketball
-   [**baseballr**](https://baseballr.sportsdataverse.org/) - baseball
-   [**fastRhockey**](https://fastRhockey.sportsdataverse.org/) - hockey
-   [**oddsapiR**](https://oddsapiR.sportsdataverse.org/) - betting odds
-   [**sportyR**](https://sportyR.sportsdataverse.org/) - playing surfaces
-   [**sportsdataverse-py**](https://py.sportsdataverse.org/) - the Python package
-   [**sportsdataverse-R**](https://r.sportsdataverse.org/) - the R meta-package
