---
title: "Using Statcast Pitch Data"
author: "Bill Petti, Saiem Gilani <br><a href='https://x.com/BillPetti' target='blank'><img src='https://img.shields.io/twitter/follow/BillPetti?color=blue&amp;label=%40BillPetti&amp;logo=x&amp;style=for-the-badge' alt='@BillPetti'/></a> <a href='https://github.com/BillPetti' target='blank'><img src='https://img.shields.io/github/followers/BillPetti?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@BillPetti'/></a> <a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>"
date: "`r Sys.Date()`"
output: html_document
description: "Working with pitch-by-pitch Statcast data from Baseball Savant - acquiring it with <code>statcast_search()</code> and shaping it for pitch-level analysis."
opengraph:
  image:
    src: "https://raw.githubusercontent.com/sportsdataverse/baseballr-data/main/themes/baseballr_gh.png"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE,
  out.width = "100%",
  dpi = 320
)
```

In this example, the `baseballr` package is used to acquire Statcast data for Corbin Burnes for the 2021 season.

This data is then used to generate different plots to showcase his arsenal and his cutter's velocity by inning.

# Load Packages

```{r load-packages}
library(baseballr)
library(dplyr)
library(ggplot2)
```


# Find Corbin Burnes' MLBAM ID

```{r get-id}
burnes_id <- baseballr::playerid_lookup(last_name = "Burnes", first_name = "Corbin") %>% 
  dplyr::pull(mlbam_id)
```

# Use Burnes' ID To Load Statcast Data

```{r load-statcast}
burnes_data <- baseballr::statcast_search_pitchers(start_date = "2021-03-01",
                                                   end_date = "2021-12-01",
                                                   pitcherid = burnes_id)
```
This block will download all of Corbin Burnes' pitches from March 1st through December 1st.

# Clean Data

If this is your first time looking at Statcast data I recommend looking at their [documentation](https://baseballsavant.mlb.com/csv-docs) for for the dataset returned. It'll walk you through the data represented by each column and give you a better idea of the data points collected for each pitch.

Some of the more common data points used for pitching analysis:

- `pitcher` : Pitcher's MLBAM ID
- `pitch_name`/`pitch_type` : Pitch Name/Type
- `release_speed` : Velocity
- `pfx_z`/`pfx_x` : Pitch Movement
- `release_spin_rate` : Spin Rate
- `spin_axis` : Spin Axis
- `release_pos_z`/`release_pos_z`/`extension` : Release Point
- `plate_z`/`plate_x` : Pitch Location


Since we're going to be making a scatterplot of Corbin Burnes' pitch movement, we need to make sure we have the data in the proper format to match a traditional movement plot. The `pfx_x` and `pfx_z` columns are both in feet so let's create two new columns and convert them to inches. `pfx_x` is also from the catcher's point of view so let's also reverse it to be from the pitcher's.

```{r clean-data}
# The glimpse function is something I use regularly
# to quickly preview the data I'm working with.
# Try it out if you haven't used it before!
# 
# 
# burnes_data %>% dplyr::glimpse()

burnes_cleaned_data <- burnes_data %>% 
  # Only keep rows with pitch movement readings
  # and during the regular season
  dplyr::filter(!is.na(pfx_x), !is.na(pfx_z),
                game_type == "R") %>% 
  dplyr::mutate(pfx_x_in_pv = -12*pfx_x,
                pfx_z_in = 12*pfx_z)
  
```

# Create A Movement Plot

Now that we've created our new columns, let's use them to plot how Corbin Burnes' pitches move. 

```{r movement-plot}
# Make a named vector to scale pitch colors with
pitch_colors <- c("4-Seam Fastball" = "red",
                  "2-Seam Fastball" = "blue",
                  "Sinker" = "cyan",
                  "Cutter" = "violet",
                  "Fastball" = "black",
                  "Curveball" = "green",
                  "Knuckle Curve" = "pink",
                  "Slider" = "orange",
                  "Changeup" = "gray50",
                  "Split-Finger" = "beige",
                  "Knuckleball" = "gold")

# Find unique pitch types to not have unnecessary pitches in legend
burnes_pitch_types <- unique(burnes_cleaned_data$pitch_name)

burnes_cleaned_data %>% 
  ggplot2::ggplot(ggplot2::aes(x = pfx_x_in_pv, y = pfx_z_in, color = pitch_name)) +
  ggplot2::geom_vline(xintercept = 0) +
  ggplot2::geom_hline(yintercept = 0) +
  # Make the points slightly transparent
  ggplot2::geom_point(size = 1.5, alpha = 0.25) +
  # Scale the pitch colors to match what we defined above
  # and limit it to only the pitches Burnes throws
  ggplot2::scale_color_manual(values = pitch_colors,
                              limits = burnes_pitch_types) +
  # Scale axes and add " to end of labels to denote inches
  ggplot2::scale_x_continuous(limits = c(-25,25),
                              breaks = seq(-20,20, 5),
                              labels = scales::number_format(suffix = "\"")) +
  ggplot2::scale_y_continuous(limits = c(-25,25),
                              breaks = seq(-20,20, 5),
                              labels = scales::number_format(suffix = "\"")) +
  ggplot2::coord_equal() +
  ggplot2::labs(title = "Corbin Burnes Pitch Movement",
                subtitle = "2021 MLB Season | Pitcher's POV",
                caption = "Data: Baseball Savant via baseballr", 
                x = "Horizontal Break",
                y = "Induced Vertical Break",
                color = "Pitch Name")
```

I like to use `pitch_name` to color my scatterplots as it gives the full pitch name in the legend, but `pitch_type` would also work if you prefer the shorter abbreviations (ex: 4-Seam Fastball = FF). If you were to use `pitch_type` instead, be sure to make a new vector for the colors.

# Velo By Inning

Now let's take a look at Corbin Burnes' pitch velo by inning for his trademark cutter.

```{r velo-by-inning}

burnes_velo_by_inning <- burnes_cleaned_data %>% 
  dplyr::filter(pitch_name == "Cutter") %>% 
  dplyr::group_by(inning, pitch_name) %>% 
  dplyr::summarize(average_velo = mean(release_speed, na.rm = TRUE))

burnes_velo_by_inning %>% 
  ggplot2::ggplot(ggplot2::aes(x = inning, y = average_velo, color = pitch_name)) +
  ggplot2::geom_line(size = 1.5, alpha = 0.5, show.legend = FALSE) +
  ggplot2::geom_point(size = 3, show.legend = FALSE) +
  ggplot2::scale_color_manual(values = pitch_colors) +
  ggplot2::scale_x_continuous(breaks = 1:9) +
  ggplot2::scale_y_continuous(limits = c(90, 100)) +
  ggplot2::labs(title = "Corbin Burnes Cutter Velo By Inning",
                subtitle = "2021 MLB Season",
                caption = "Data: Baseball Savant via baseballr",
                x = "Inning",
                y = "Average Velo")

```


# Conclusion

These were just two examples of things you can do with pitch data acquired using baseballr. Statcast data goes back to 2015 and contains a multitude of data points for each pitch/batted ball event so there's endless things to go research! 



# Beyond MLB: minors and the World Baseball Classic

Baseball Savant serves minor-league and World Baseball Classic Statcast data
from separate search routes, so those games never appear in `statcast_search()`
results. The package wraps them with the same interface and columns:

```{r, eval=FALSE}
# Triple-A pitch data (Statcast tracking in MiLB parks began 2023)
milb <- statcast_search_minors(start_date = "2024-06-01", end_date = "2024-06-01")

# World Baseball Classic
wbc <- statcast_search_wbc(start_date = "2023-03-21", end_date = "2023-03-21")
```

# **Our Authors**
-   [Bill Petti](https://x.com/BillPetti)
<a href='https://x.com/BillPetti' target='blank'><img src='https://img.shields.io/twitter/follow/BillPetti?color=blue&amp;label=%40BillPetti&amp;logo=x&amp;style=for-the-badge' alt='@BillPetti'/></a>
<a href='https://github.com/BillPetti' target='blank'><img src='https://img.shields.io/github/followers/BillPetti?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@BillPetti'/></a>
-   [Saiem Gilani](https://x.com/saiemgilani)
<a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a>
<a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>

## **Our Contributors**
-   [Ben Baumer](https://x.com/BaumerBen)
<a href='https://x.com/BaumerBen' target='blank'><img src='https://img.shields.io/twitter/follow/BaumerBen?color=blue&amp;label=%40BaumerBen&amp;logo=x&amp;style=for-the-badge' alt='@BaumerBen'/></a>
<a href='https://github.com/beanumber' target='blank'><img src='https://img.shields.io/github/followers/beanumber?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@beanumber'/></a>
-   [Ben Dilday](https://x.com/BenDilday)
<a href='https://x.com/BenDilday' target='blank'><img src='https://img.shields.io/twitter/follow/BenDilday?color=blue&amp;label=%40BenDilday&amp;logo=x&amp;style=for-the-badge' alt='@BenDilday'/></a>
<a href='https://github.com/bdilday' target='blank'><img src='https://img.shields.io/github/followers/bdilday?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@bdilday'/></a>
-   [Robert Frey](https://x.com/RobertFrey40)
<a href='https://x.com/RobertFrey40' target='blank'><img src='https://img.shields.io/twitter/follow/RobertFrey40?color=blue&amp;label=%40RobertFrey40&amp;logo=x&amp;style=for-the-badge' alt='@RobertFrey40'/></a>
<a href='https://github.com/robert-frey' target='blank'><img src='https://img.shields.io/github/followers/robert-frey?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@robert-frey'/></a>
-   [Camden Kay](https://x.com/k_camden)
<a href='https://x.com/k_camden' target='blank'><img src='https://img.shields.io/twitter/follow/k_camden?color=blue&amp;label=%40k_camden&amp;logo=x&amp;style=for-the-badge' alt='@k_camden'/></a>
<a href='https://github.com/camdenk' target='blank'><img src='https://img.shields.io/github/followers/camdenk?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@camdenk'/></a>

## **Citation**

To cite the [**`baseballr`**](https://baseballr.sportsdataverse.org/) R package in publications, use:

BibTeX Citation

```bibtex
@misc{baseballr,
  author = {Bill Petti and Saiem Gilani},
  title = {baseballr: An R Package for Baseball Data Acquisition and Analysis},
  url = {https://baseballr.sportsdataverse.org/},
  year = {2026}
}
```

## **Related SportsDataverse packages**

-   [**cfbfastR**](https://cfbfastR.sportsdataverse.org/) - college football
-   [**hoopR**](https://hoopR.sportsdataverse.org/) - men's basketball
-   [**wehoop**](https://wehoop.sportsdataverse.org/) - women's basketball
-   [**baseballr**](https://baseballr.sportsdataverse.org/) - baseball
-   [**fastRhockey**](https://fastRhockey.sportsdataverse.org/) - hockey
-   [**oddsapiR**](https://oddsapiR.sportsdataverse.org/) - betting odds
-   [**sportyR**](https://sportyR.sportsdataverse.org/) - playing surfaces
-   [**sportsdataverse-py**](https://py.sportsdataverse.org/) - the Python package
-   [**sportsdataverse-R**](https://r.sportsdataverse.org/) - the R meta-package
